---
Auther: "Vedant Shah 987144"
title: "ABDA_Week1 Vedant Shah 987144"
output: html_document
---


```{r, warning=FALSE}

library(tidyverse)
library(ggbeeswarm)
library(aida)

dolphin <- aida::aidata

```



### HOMEWORK Data Wrangling###

### Exercise 2 

(a) Take the dolphin data set and store a reduced variant of it. The new data frame should contain only the columns `condition`, `group`, and `xpos_flips`, `correct`. And within the `correct` vector, we are only interested in the correct trials (= 1). Filter accordingly.

```{r x2a}
  
dolphin_selected <- dolphin %>%
  dplyr::select(condition, group, xpos_flips, correct) %>% filter(correct == 1)
head(dolphin_selected)

```

(b) Create an aggregated data frame that contains the mean `xpos_flips` value and the standard deviation for `group` and `condition`.

```{r x2b}
dolphin_aggregate <-
  dolphin_selected %>%
  group_by(condition, group) %>%
  summarise(mean_flips = mean(xpos_flips), 
            sd_flips  = sd(xpos_flips))

dolphin_aggregate
```

(c) Rename the new vectors for the mean xflips and their sd to `xflips_mean` and `xflips_sd`

```{r}
dolphin_aggregate <- dolphin_aggregate %>% rename(xflips_mean = mean_flips, xflips_sd = sd_flips)
dolphin_aggregate
```

(d) Do all of the above (a-c) in one pipeline.

```{r}
dolphin_pipeline <- dolphin %>%
  dplyr::select(condition, group, xpos_flips, correct) %>% filter(correct == 1) %>%
group_by(condition, group) %>%
  summarise(mean_flips = mean(xpos_flips), 
            sd_flips  = sd(xpos_flips)) %>% rename(xflips_mean = mean_flips, xflips_sd = sd_flips)
  
dolphin_pipeline

```

### HOMEWORK Data Viz ###

## Exercise 2

a) Plot AUC values as a function of group in a density plot (`geom_density`).

b) Color code the curves (the border and the filling) as follows: "orange" for the click group and "black" for the touch group.

c) Make the density curves semi-transparent with the `alpha` argument

d) Add the aida_theme to the plot.

e) BONUS: Add the mean values for both groups into the density plot as a line.

```{r x2}


mean <- dolphin %>% group_by(group) %>% summarise(mean = mean(AUC))

ggplot(dolphin, aes(x= AUC, fill = group, color = group)) +
  geom_density(alpha = 0.5) +
  ggtitle("a density plot") +
  scale_fill_manual(
                    guide = guide_legend("Group"),
                    values = c("orange", "black")) +
    scale_colour_manual(
                    guide = guide_legend("Group"),
                    values = c("orange", "black")) +
  geom_vline(data = mean, aes(xintercept = mean, color = group))+
  theme_aida()

```


